---
title: "InvaCost Spatial Extrapolation"
author: "Gabriel Caetano, Ivan Jaric, Franck Courchamp"
date: "`r Sys.Date()`"
output: 
  pdf_document: default
  rmarkdown::html_vignette:
  toc: true
vignette: >
  %\VignetteIndexEntry{pcpi} 
  %\VignetteEngine{knitr::rmarkdown} 
  %\VignetteEncoding{UTF-8}
---

# To do

- Include installation of all packages
- Find better records of invasive species distribution
- Investigate NAs for Maldives and Namibia
- Investigate NAs for Phyllum
- download GBIF records again

# Set up

Install packages (turned off automatic evaluation so it doesn't try to install them every time)

```{r, eval = FALSE}
install.packages("magrittr")
install.packages("stringr")
install.packages("rgbif")
install.packages("pbapply")
install.packages("caret")
install.packages("forcats")
install.packages("wbstats")
```

Load packages

```{r}
library(magrittr)
library(rgbif)
library(ggplot2)
```

Load data

```{r}
inva_db <-
  invacost::getInvaCostVersion()
```

# Data Exploration

What does the database looks like?

```{r}
head(inva_db)
```

Remove low reliability data points, get only data points at country level, and also only data points with unique species, country and sector

```{r}
inva_db_clean <- 
  inva_db[inva_db$Method_reliability == "High" & 
            inva_db$Spatial_scale == "Country" &
            inva_db$Implementation == "Observed" &
            !stringr::str_detect(inva_db$Official_country, "/"),]

inva_db_clean$Impacted_sector[stringr::str_detect(inva_db_clean$Impacted_sector, "/")] <- "Mixed"
inva_db_clean$Phylum[stringr::str_detect(inva_db_clean$Phylum, "/")] <- "Mixed"
inva_db_clean$Class[stringr::str_detect(inva_db_clean$Class, "/")] <- "Mixed"
inva_db_clean$Geographic_region[stringr::str_detect(inva_db_clean$Geographic_region, "/")] <- "Mixed"
inva_db_clean$Environment[stringr::str_detect(inva_db_clean$Environment, "/")] <- "Mixed"
```

Fix some typos and remove unspecified countries and impact sectors

```{r}
inva_db_clean$Impacted_sector[inva_db_clean$Impacted_sector == "Forestry " ] <- "Forestry"
inva_db_clean$Impacted_sector[inva_db_clean$Impacted_sector == "Foresty" ] <- "Forestry"
inva_db_clean$Impacted_sector[inva_db_clean$Impacted_sector == "Public and social welfare" ] <- "Public and Social Welfare"

inva_db_clean$Official_country[inva_db_clean$Official_country == "Kanya" ] <- "Kenya"
inva_db_clean$Official_country[inva_db_clean$Official_country == "Republic of Serbia" ] <- "Serbia"
inva_db_clean$Official_country[inva_db_clean$Official_country == "The Bahamas" ] <- "Bahamas"
inva_db_clean$Official_country[inva_db_clean$Official_country == "United Republic of Tanzania" ] <- "Tanzania"
inva_db_clean$Official_country[inva_db_clean$Official_country == "United States of America"] <- "United States"
inva_db_clean$Official_country[inva_db_clean$Official_country == "Venezuela"] <- "Venezuela, Bolivarian Republic of"
inva_db_clean$Official_country[inva_db_clean$Official_country == "Vietnam"] <- "Viet Nam"
inva_db_clean$Official_country[inva_db_clean$Official_country == "Tanzania"] <- "Tanzania, United Republic of"
inva_db_clean$Official_country[inva_db_clean$Official_country == "South Korea"] <- "Korea, Republic of"
inva_db_clean$Official_country[inva_db_clean$Official_country == "Russia"] <- "Russian Federation"
inva_db_clean$Official_country[inva_db_clean$Official_country == "Brunei"] <- "Brunei Darussalam"
inva_db_clean$Official_country[inva_db_clean$Official_country == "Laos"] <- "Lao People's Democratic Republic"
inva_db_clean$Official_country[inva_db_clean$Official_country == "East Timor"] <- "Timor-Leste"
inva_db_clean$Official_country[inva_db_clean$Official_country == "Bolivia"] <- "Bolivia, Plurinational State of"
inva_db_clean$Official_country[inva_db_clean$Official_country == "Ivory Coast"] <- "Cote d'Ivoire"
inva_db_clean$Official_country[inva_db_clean$Official_country == "Macedonia"] <- "Macedonia, the former Yugoslav Republic of"
inva_db_clean$Official_country[inva_db_clean$Official_country == "Moldova"] <- "Moldova, Republic of"
inva_db_clean$Official_country[inva_db_clean$Official_country == "Democratic Republic of the Congo"] <- "Congo, the Democratic Republic of the"
inva_db_clean$Official_country[inva_db_clean$Official_country == "Iran"] <- "Iran, Islamic Republic of"
inva_db_clean$Official_country[inva_db_clean$Official_country == "Federal States of Micronesia"] <- "Micronesia, Federated States of"
inva_db_clean$Official_country[inva_db_clean$Official_country == "Czechia"] <- "Czech Republic"

inva_db_clean <- inva_db_clean[!inva_db_clean$Impacted_sector %in% c("Unspecified"),]
inva_db_clean <- inva_db_clean[!inva_db_clean$Official_country %in% c("31 countries",
                                                                      "African Countries",
                                                                      "Diverse",
                                                                      "EU27+",
                                                                      "Unspecified"),]

inva_db_clean <- inva_db_clean[!duplicated(inva_db_clean),]

dim(inva_db_clean)

write.csv(inva_db_clean, "inva_db_clean.csv", row.names = FALSE)
```

It seems the there are a few data points with multiple species, countries and impacted sectors. Let's create a new table with only unique combinations.

```{r}
scs_df <-
  data.frame(
    Species = inva_db_clean$Species,
    Official_country = inva_db_clean$Official_country,
    Impacted_sector  = inva_db_clean$Impacted_sector)

scs_df <- scs_df[complete.cases(scs_df),]

scs_unq <-
  lapply(1:nrow(scs_df), function(r){
    
    row <- scs_df[r,]
    #print(row.names(row))
    row.names(row) <- NULL
    
    which_multi <- stringr::str_detect(row, "/")
    
    if(!any(which_multi)) return(row)
    
    split_cases <-
      sapply(row[which_multi], stringr::str_split, "/")
    
    out <-
      do.call(expand.grid, split_cases) %>% 
      data.frame(row[!which_multi] )
    
    out[match(colnames(out), names(row))]
    
  }) %>% 
  do.call(rbind, .)

scs_unq <- 
  scs_unq[!stringr::str_detect(scs_unq$Species, "\\."),]

dim(scs_unq)
```

# Completeness analysis

Look up species names in GBIF taxonomy (this may take a while)

```{r, eval = FALSE}
gbif_lookup <- 
  pbapply::pblapply(unique(scs_unq$Species), 
                    rgbif::name_lookup, 
                    verbose = FALSE)

saveRDS(gbif_lookup, "gbif_lookup.rds")
```

See which names found no match in GBIF

```{r}
gbif_lookup <- readRDS("gbif_lookup.rds")

name_match <- 
  sapply(1:length(unique(scs_unq$Species)), 
         function(i){
           
           unique(scs_unq$Species)[i] %in%
             c(gbif_lookup[[i]]$data$species, 
               gbif_lookup[[i]]$data$scientificName)
           
         })

unique(scs_unq$Species)[!name_match]
```

Fix mispellings

```{r}
# Mispellings
scs_unq$Species[scs_unq$Species == "Felix catus"] <- "Felis catus"
scs_unq$Species[scs_unq$Species == "Rarrus rattus"] <- "Rattus rattus"
scs_unq$Species[scs_unq$Species == "italicum"] <- "Arum italicum"
scs_unq$Species[scs_unq$Species == "Cylindroptunia sp."] <- "Cylindropuntia sp."
scs_unq$Species[scs_unq$Species == "Bos taurus "] <- "Bos taurus"
scs_unq$Species[scs_unq$Species == "Andean ?otato latent tymovirus"] <- "Andean potato latent tymovirus"
scs_unq$Species[scs_unq$Species == " Proposis juliflora"] <- "Proposis juliflora"
scs_unq$Species[scs_unq$Species == " Proposis velutina"] <- "Proposis velutina"
scs_unq$Species[scs_unq$Species == " Solenopsis richteri"] <- "Solenopsis richteri"
scs_unq$Species[scs_unq$Species == "Gaeumannomyces graminis\x86var.\x86tritici"] <- "Gaeumannomyces graminis var. tritici"
scs_unq$Species[scs_unq$Species == "Spartina\xa0spp."] <- "Spartina spp."
scs_unq$Species[scs_unq$Species == "Candidatus\xa0liberibacter"] <- "Candidatus liberibacter"
scs_unq$Species[scs_unq$Species == "Casuarina equisetifolia "] <- "Casuarina equisetifolia"
scs_unq$Species[scs_unq$Species == "Tilletia Indica"] <- "Tilletia indica"
scs_unq$Species[scs_unq$Species == "Ralstonia Solanacearum"] <- "Ralstonia solanacearum"
scs_unq$Species[scs_unq$Species == "Pacifastacus Leniusculus"] <- "Pacifastacus leniusculus"
scs_unq$Species[scs_unq$Species == "Opuntia ficus-indica "] <- "Opuntia ficus-indica"
scs_unq$Species[scs_unq$Species == "Blatella germanica"] <- "Blattella germanica"
scs_unq$Species[scs_unq$Species == "Peach  rosette  mosaic  nepovirus "] <- "Peach  rosette  mosaic  nepovirus"
scs_unq$Species[scs_unq$Species == "Oligonychus metasequoia"] <- "Oligonychus metasequoiae" 
scs_unq$Species[scs_unq$Species == "Proposis pallida"] <- "Prosopis pallida"
scs_unq$Species[scs_unq$Species == "Proposis gladulosa"] <- "Prosopis glandulosa"
scs_unq$Species[scs_unq$Species == "Proposis juliflora"] <- "Prosopis juliflora"
scs_unq$Species[scs_unq$Species == "Proposis velutina"] <- "Prosopis velutina" 
scs_unq$Species[scs_unq$Species == "Robinia pseudoaccia"] <- "Robinia pseudoacacia"
scs_unq$Species[scs_unq$Species == "Cercis siliquatrum"] <- "Cercis siliquastrum"
scs_unq$Species[scs_unq$Species == "Lymantria disparasiatica"] <- "Lymantria dispar asiatica"
scs_unq$Species[scs_unq$Species == "Syrindodium filiformis"] <- "Syringodium filiforme"
scs_unq$Species[scs_unq$Species == "Clerodendrum chinese"] <- "Clerodendrum chinense"
```

Look fixed list up again and get only species with exact matches in GBIF.

```{r, eval = FALSE}
gbif_lookup2 <- 
  pbapply::pblapply(unique(scs_unq$Species), 
                    rgbif::name_lookup, 
                    verbose = FALSE)

saveRDS(gbif_lookup2, "gbif_lookup2.rds")
```

```{r}
gbif_lookup2 <- readRDS("gbif_lookup2.rds")

name_match <- 
  sapply(1:length(unique(scs_unq$Species)), 
         function(i){
           
           unique(scs_unq$Species)[i] %in%
             c(gbif_lookup2[[i]]$data$species, 
               gbif_lookup2[[i]]$data$scientificName)
           
         })

unique(scs_unq$Species)[!name_match]

scs_unq_gbif <- 
  scs_unq[scs_unq$Species %in% unique(scs_unq$Species)[name_match],]

scs_unq_gbif$iso2 <- 
  rgbif::isocodes$code[match(scs_unq_gbif$Official_country,
                             rgbif::isocodes$name)]

scs_unq_gbif$iso2[scs_unq_gbif$Official_country == "Kosovo"] <- "XK"

scs_unq_gbif$Geographic_region <- 
  inva_db_clean$Geographic_region[match(scs_unq_gbif$Official_country,
                                        inva_db_clean$Official_country)]

country_region <-
  scs_unq_gbif[c(2,4,5)] %>% 
  unique()

write.csv(country_region, "country_region.csv", row.names = FALSE)
```

Get GBIF taxon keys

```{r}
gbif_taxon_keys <- 
  unique(scs_unq_gbif$Species) %>% 
  rgbif::name_backbone_checklist() %>% 
  dplyr::filter(!matchType == "NONE") %>%
  dplyr::pull(usageKey) 
```

List country of occurrence for each species

```{r, warning = FALSE, eval = FALSE}
gbif_taxon_key <- 
  unique(scs_unq_gbif$Species) %>% 
  rgbif::name_backbone_checklist() %>% 
  dplyr::filter(!matchType == "NONE") %>%
  dplyr::pull(usageKey) 

pbapply::pblapply(1:length(gbif_taxon_key), function(i){
  
  countries <- 
    rgbif::count_facet(gbif_taxon_key[i], 
                       by = "country", 
                       countries = 249)
  
  if(sum(countries$count) == 0) return(NULL)
  
  out <-
    data.frame(sp = unique(scs_unq_gbif$Species)[i], 
               country = countries$country[countries$count > 0])
  
  write.csv(out, paste0("country_sp_", i, ".csv"), row.names = FALSE)
  
})

country_species <-
  pbapply::pblapply(list.files(pattern = "^country_sp_", 
                               full.names = TRUE), 
                    function(fl){
                      
                      read.csv(fl)
                      
                    }) %>% 
  do.call(rbind, .)

country_species <- country_species[complete.cases(country_species),]
scs_unq_gbif <- scs_unq_gbif[complete.cases(scs_unq_gbif),]

saveRDS(country_species, "country_species.rds")
```

Calculate proportion of sampled species per country

```{r}
country_species <- readRDS("country_species.rds")

completeness_df <-
  lapply(unique(scs_unq_gbif$iso2), 
         function(cntr){
           
           inva_sp <-
             scs_unq_gbif$Species[scs_unq_gbif$iso2 == cntr] %>% 
             unique
           
           cntr_sp <-
             country_species$sp[country_species$country == cntr] %>% 
             unique
           
           sp_prop <- sum(inva_sp %in% cntr_sp)/length(cntr_sp)
           
           data.frame(country = cntr,
                      sp_prop = sp_prop)
           
         }) %>% 
  do.call(rbind, .)

write.csv(completeness_df, "completeness_df.csv", row.names = FALSE)
```

Plot countries with more than 1% data completeness

```{r}
completeness_df[completeness_df$sp_prop > 0.01,] %>% 
  ggplot(aes(x = forcats::fct_rev(forcats::fct_reorder(country, sp_prop)),
             y = sp_prop,
             fill = sp_prop)) +
  geom_col() +
  theme_classic()
```

Separate data by impacted sector

```{r}
completeness_sector_df <-
  lapply(unique(scs_unq_gbif$Impacted_sector), function(sect){
    
    scs_unq_sect <- 
      scs_unq_gbif[scs_unq_gbif$Impacted_sector == sect,]
    
    country_species_sect <- 
      country_species[country_species$sp %in% scs_unq_sect$Species,]
    
    completeness_df_this_sect <-
      lapply(unique(scs_unq_sect$iso2), 
             function(cntr){
               
               inva_sp <-
                 scs_unq_sect$Species[scs_unq_sect$iso2 == cntr] %>% 
                 unique
               
               cntr_sp <-
                 country_species_sect$sp[country_species_sect$country == cntr] %>% 
                 unique
               
               sp_prop <- sum(inva_sp %in% cntr_sp)/length(cntr_sp)
               
               data.frame(country = cntr,
                          sp_prop = sp_prop)
               
             }) %>% 
      do.call(rbind, .)
    
    completeness_df_this_sect$sector <- sect
    
    completeness_df_this_sect
    
  }) %>% 
  do.call(rbind, .)
```

Plot countries with more than 10% data completeness for each impacted sector

```{r}
completeness_sector_df[completeness_sector_df$sector == "Agriculture" &
                         completeness_sector_df$sp_prop > 0.01,] %>% 
  ggplot(aes(x = forcats::fct_rev(forcats::fct_reorder(country, sp_prop)),
             y = sp_prop,
             fill = sp_prop)) +
  geom_col() +
  theme_classic()

completeness_sector_df[completeness_sector_df$sector == "Environment" &
                         completeness_sector_df$sp_prop > 0.01,] %>% 
  ggplot(aes(x = forcats::fct_rev(forcats::fct_reorder(country, sp_prop)),
             y = sp_prop,
             fill = sp_prop)) +
  geom_col() +
  theme_classic()

completeness_sector_df[completeness_sector_df$sector == "Authorities-Stakeholders" &
                         completeness_sector_df$sp_prop > 0.01,] %>% 
  ggplot(aes(x = forcats::fct_rev(forcats::fct_reorder(country, sp_prop)),
             y = sp_prop,
             fill = sp_prop)) +
  geom_col() +
  theme_classic()

completeness_sector_df[completeness_sector_df$sector == "Public and Social Welfare" &
                         completeness_sector_df$sp_prop > 0.01,] %>% 
  ggplot(aes(x = forcats::fct_rev(forcats::fct_reorder(country, sp_prop)),
             y = sp_prop,
             fill = sp_prop)) +
  geom_col() +
  theme_classic()

completeness_sector_df[completeness_sector_df$sector == "Health" &
                         completeness_sector_df$sp_prop > 0.01,] %>% 
  ggplot(aes(x = forcats::fct_rev(forcats::fct_reorder(country, sp_prop)),
             y = sp_prop,
             fill = sp_prop)) +
  geom_col() +
  theme_classic()

completeness_sector_df[completeness_sector_df$sector == "Forestry" &
                         completeness_sector_df$sp_prop > 0.01,] %>% 
  ggplot(aes(x = forcats::fct_rev(forcats::fct_reorder(country, sp_prop)),
             y = sp_prop,
             fill = sp_prop)) +
  geom_col() +
  theme_classic()

completeness_sector_df[completeness_sector_df$sector == "Fishery" &
                         completeness_sector_df$sp_prop > 0.01,] %>% 
  ggplot(aes(x = forcats::fct_rev(forcats::fct_reorder(country, sp_prop)),
             y = sp_prop,
             fill = sp_prop)) +
  geom_col() +
  theme_classic()
```

# Cost distribution analysis

Select only phylum, sector, environment and cost

```{r}
psce_df <-
  data.frame(
    Phylum = inva_db_clean$Phylum,
    Class = inva_db_clean$Class,
    Geographic_region = inva_db_clean$Geographic_region,
    Sector  = inva_db_clean$Impacted_sector,
    Environment = inva_db_clean$Environment,
    Cost = inva_db_clean$Cost_estimate_per_year_2017_USD_exchange_rate,
    Country = inva_db_clean$Official_country,
    Type_of_cost = inva_db_clean$Type_of_cost_merged)

psce_unq <- psce_df[complete.cases(psce_df),]

dim(psce_unq)
```

Group Phylla into broader groups

```{r}
psce_unq$Phylum %>% unique

psce_unq$Group <- NA
psce_unq$Group[psce_unq$Phylum == "Chordata"] <- "Vertebrates"

psce_unq$Group[psce_unq$Phylum == "Arthropoda"] <- "Arthropods"

psce_unq$Group[psce_unq$Phylum == "Nematoda"] <- "Invertebrates"          
psce_unq$Group[psce_unq$Phylum == "Platyhelminthes"] <- "Invertebrates"
psce_unq$Group[psce_unq$Phylum == "Annelida"] <- "Invertebrates"
psce_unq$Group[psce_unq$Phylum == "Cnidaria"] <- "Invertebrates"
psce_unq$Group[psce_unq$Phylum == "Mollusca"] <- "Invertebrates"  

psce_unq$Group[psce_unq$Phylum == "Tracheophyta"] <- "Plants" 
psce_unq$Group[psce_unq$Phylum == "Ochrophyta"] <- "Plants"
psce_unq$Group[psce_unq$Phylum == "Chlorophyta"] <- "Plants"

psce_unq$Group[psce_unq$Phylum == "Oomycota"] <- "Fungi"     
psce_unq$Group[psce_unq$Phylum == "Basidiomycota"] <- "Fungi"
psce_unq$Group[psce_unq$Phylum == "Basidiomycota"] <- "Fungi"
psce_unq$Group[psce_unq$Phylum == "Chytridiomycota"] <- "Fungi"
psce_unq$Group[psce_unq$Phylum == "Ascomycota"] <- "Fungi"
psce_unq$Group[psce_unq$Phylum == "Ascomyceta"] <- "Fungi"

psce_unq$Group[psce_unq$Phylum == "Proteobacteria"] <- "Microbes" 
psce_unq$Group[psce_unq$Phylum == "Cressdnaviricota"] <- "Microbes"            
psce_unq$Group[psce_unq$Phylum == "Negarnaviricota"] <- "Microbes"   
psce_unq$Group[psce_unq$Phylum == "Pisuviricota"] <- "Microbes"       
psce_unq$Group[psce_unq$Phylum == "Kitrinoviricota"] <- "Microbes"    
psce_unq$Group[psce_unq$Phylum == "Nucleocytoviricota"] <- "Microbes"           
psce_unq$Group[psce_unq$Phylum == "Peploviricota"] <- "Microbes" 
```

Plot cost distribution by group and sector, for each environment

```{r}
psce_unq %>% 
  ggplot(aes(x = Sector, 
             y = log(Cost))) +
  geom_boxplot() + 
  theme_classic()

psce_unq %>% 
  ggplot(aes(x = Group, 
             y = log(Cost))) +
  geom_boxplot() + 
  theme_classic()

psce_unq %>% 
  ggplot(aes(x = Sector, 
             y = log(Cost),
             fill = Group)) +
  geom_boxplot() + 
  theme_classic()

psce_unq[psce_unq$Environment == "Aquatic",] %>% 
  ggplot(aes(x = Sector, 
             y = log(Cost),
             fill = Group)) +
  geom_boxplot() + 
  theme_classic()

psce_unq[psce_unq$Environment == "Terrestrial",] %>% 
  ggplot(aes(x = Sector, 
             y = log(Cost),
             fill = Group)) +
  geom_boxplot() + 
  theme_classic()

write.csv(psce_unq, "psce_unq.csv", row.names = FALSE)
```

# Get country data from WorldBank

Find which indicators are available for what countries

```{r}
cachelist <- wbstats::wb_cache()
```

Download all indicators from WorldBank and extract country information

```{r, eval = FALSE, warning = FALSE}
to_download <- 1:length(cachelist$indicators$indicator_id)

# In case downloading is interrupted:
all_files <- 1:length(cachelist$indicators$indicator_id)
downloaded_files <-
  list.files(pattern = "^indicator_") %>%
  stringr::str_sub(11, -5) %>%
  as.numeric()
to_download <- all_files[!all_files %in% downloaded_files]

inds_cntr <- 
  pbapply::pblapply(to_download, 
                    function(i){
                      
                      #print(i)
                      ind <- cachelist$indicators$indicator_id[i]
                      
                      this_ind <- 
                        try(wbstats::wb_data(ind, mrnev = 1), silent = TRUE)
                      
                      if(class(this_ind)[1] == "try-error"){
                        
                        return(this_ind)
                        
                      } else if(ncol(this_ind) == 0){ 
                        
                        return(NULL)
                        
                      } else {
                        
                        saveRDS(this_ind$iso2c[!is.na(this_ind[5])],
                                paste0("indicator_", i, ".rds"))
                        
                        this_ind$iso2c[!is.na(this_ind[5])]
                        
                      }
                      
                    })
```

Read all in a single table

```{r, eval = FALSE}
inds_cntr <-
  pbapply::pblapply(list.files("C:/Users/gabri/Desktop/wb_indicators_cntr", pattern = "^indicator_", 
                               full.names = TRUE), 
                    function(fl){
                      
                      readRDS(fl)
                      
                    })

dl_index <-
  list.files("C:/Users/gabri/Desktop/wb_indicators_cntr", pattern = "^indicator_") %>%
  stringr::str_sub(11, -5) %>%
  as.numeric()

names(inds_cntr) <- cachelist$indicators$indicator_id[dl_index]

saveRDS(inds_cntr, "inds_cntr.rds")
```

Get only indicators available for all UN member states in InvaCost

```{r, eval = FALSE}
inva_cntr <- unique(scs_unq_gbif$iso2)
un_members <- read.csv("un_members.csv")
un_members$iso2 <- stringr::str_sub(un_members$iso2, 1, 2)

inva_cntr <- inva_cntr[inva_cntr %in% un_members$iso2]

compl_index <-
  sapply(inds_cntr, function(x) sum(inva_cntr %in% x) == length(inva_cntr))

compl_indicators <- names(inds_cntr)[compl_index]

complete_indicators <-
  cachelist$indicators[cachelist$indicators$indicator_id %in% compl_indicators,]

write.csv(complete_indicators[-6], 
          "complete_indicators.csv", 
          row.names = FALSE)
```

Select appropriate predictors

```{r}
complete_indicators <- 
  read.csv("complete_indicators.csv")

exclude_ind <- c(1:11, 13:21, 23, 31:64, 66, 67, 69, 71:73, 81:100, 102:105, 107, 108, 110:126, 128:133, 134:137, 138:248, 250:261, 263:274, 276, 278:462, 463, 465:477, 478:493, 495, 496, 498:515, 518, 519, 521, 522, 524:532, 535, 536, 538, 539, 541:605)

clean_indicators <- complete_indicators[-exclude_ind,]
```

Download selected indicators

```{r}
inds_clean_list <- 
  pbapply::pblapply(clean_indicators$indicator_id, 
                    function(ind){
                      
                      this_ind <- 
                        try(wbstats::wb_data(ind, mrnev = 1), silent = TRUE)
                      
                      if(class(this_ind)[1] == "try-error"){
                        
                        return(this_ind)
                        
                      } else if(ncol(this_ind) == 0){ 
                        
                        return(NULL)
                        
                      } else {
                        
                        return(this_ind)
                        
                      }
                      
                    })

saveRDS(inds_clean_list, "inds_clean_list.rds")
```

```{r}
inds_clean_list <- readRDS("inds_clean_list.rds")

errors_index <- sapply(inds_clean_list, function(x) "try-error" %in% class(x))

inds_clean_list <- inds_clean_list[!errors_index]

inds_df <-
  lapply(inds_clean_list, function(ind) ind[,c(1,5)]) %>% 
  plyr::join_all(., by='iso2c', type='full')

inds_df <- inds_df[inds_df$iso2c %in% un_members$iso2,]

inds_df <- inds_df[!apply(inds_df, 2, function(x) any(is.na(x)))]

write.csv(inds_df, "inds_df.csv", row.names = FALSE)
```

# Scrape GISD

Scrape the GISD website to find where species are native and where they are invasive

```{r, eval = FALSE}
gisd_table <-
  pbapply::pblapply(1:2099, function(i){
    
    Sys.sleep(0.5)
    
    url <- paste0("http://www.iucngisd.org/gisd/species.php?sc=", i)
    
    sp_html <-
      rvest::read_html(url)
    
    all_elements <-
      rvest::html_elements(sp_html, "div")
    
    element_id <- 
      rvest::html_attr(all_elements, "id")
    
    if("ar-col" %in% element_id & 
       "nr-col" %in% element_id){
      
      sp_name <- all_elements[which(element_id == "spe-title")] %>% rvest::html_text2()
      alien_range <- all_elements[which(element_id == "ar-col")] %>% rvest::html_text2()
      native_range <- all_elements[which(element_id == "nr-col")] %>% rvest::html_text2()
      
      split_ar <-
        stringr::str_split(alien_range, "\n")[[1]][-c(1:3)]
      
      split_nr <-
        stringr::str_split(native_range, "\n")[[1]][-c(1:3)]
      
      if(length(split_nr) == 0) split_nr <- NULL
      
      if(length(split_ar) > 0){
        
        split_ar_fixed <-
          sapply(1:length(split_ar), function(r){
            
            stringr::str_sub(split_ar[r], stringr::str_locate(split_ar, "] ")[r,2] + 1)
            
          })
        
      } else {split_ar_fixed <- NULL}
      
      
      data.frame(sp = sp_name, 
                 country = c(split_ar_fixed, split_nr), 
                 status = c(rep("invasive", length(split_ar_fixed)),
                            rep("native", length(split_nr))))
      
    } else { NULL }
    
  }) %>% 
  do.call(rbind, .)

write.csv(gisd_table, "gisd_table.csv")
```

See how many of the species in InvaCost filtered have data in GISD

```{r}
gisd_table <- read.csv("gisd_table.csv")

sp_inva <- 
  inva_db_clean$Species[!stringr::str_detect(inva_db_clean$Species, "\\.")] %>% 
  unique

sum(sp_inva %in% gisd_table$sp)/length(sp_inva)
```
